---
title: "Comparing Two Data File"
output: html_document
---
---
title: "Comparing Years Over Place and GIS Data Files"
output: html_document
---

Author: Rachel Chung

The purpose of this script is to check whether the Years Over Place Settlement R.Data file and GIS Google Sheet Data File have the same name settlement names and total instances. 

Importing packages and data

```{r}
library(descr)
#load GIS file
#load Years Over Place file
GIS_Jesse <- Years_Place_Settlements_GIS_06_2020_1_
```


Checking if the settlement names and number of settlements are the same

```{r setup, include=FALSE}
#finding duplicates in place names 
GIS_Jesse$`Chronicle Name` [duplicated(GIS_Jesse$`Chronicle Name` )]
wide_settle$Version_1_Names[duplicated(wide_settle$Version_1_Names)]

#subsetting based on place names to check both datasets have the same list of names 
nrow(GIS_Jesse)
nrow(wide_settle)
subset(wide_settle, !(Version_1_Names %in% GIS_Jesse$`Chronicle Name`)) 
subset(GIS_Jesse, !(`Chronicle Name` %in% wide_settle$Version_1_Names))
```

Checking if the total instances are the same (there are 2 methods)

Method 1: provides an basic understanding if the total for each row match up in both datasets 

```{r}
names(GIS_Jesse)[20]<-"GIS_TI"

#merging both dataset using their row names 
checking_TI <- merge(by.x = "Chronicle Name", by.y = "Version_1_Names" ,GIS_Jesse[,c(1,20)],wide_settle[,c(1,5)])

#creating an column that will tell us if the total instances column is the same in both datasets
checking_TI$Total_Instance_Inconsistencies <- checking_TI$GIS_TI == checking_TI$Total_Instances 
```


Method 2: providing an exact tool for comparing mentions. The following loop compares each cell (an cell meaning total mentions in an particular year for an particular row) against an equivalent cell in the other dataset 

Cell-by-cell comparison

```{r}
#wide_settle <- wide_settle[-c(106, 107, 273,294), ]
year_inconsistency <- as.data.frame(wide_settle[,1])
names(year_inconsistency)[1] <- "Item"
year_inconsistency$Inconsistent_Years <- NA

GIS_Jesse [is.na(GIS_Jesse )] <- 0 #converting all NAs to 0s 

#must sort the data name by row names in order for the loop to work properly - otherwise the data could be comparing the incorrect cells in the data
wide_settle <- wide_settle[order(wide_settle$Version_1_Names),] 

GIS_Jesse <- GIS_Jesse[order(GIS_Jesse$`Chronicle Name`),]

for (i in 1:nrow(wide_settle)){ #the loop iterating cell by cell 
  for (c in 6:ncol(wide_settle)){
    part1 = wide_settle[i,c]
    part2 = as.data.frame(GIS_Jesse[i,c+15])
    result = part1== part2
    if (result == FALSE){ #if two cells have different total instances, R will print the inconsistent values as the program runs 
      cat(wide_settle[i,1], names(wide_settle[c]), " --> ")
      cat("Rachel Data Value:", part1, " | Jesse Data Value: ", part2[,1], "\n")
      year_inconsistency[i,2] <- paste(year_inconsistency[i,2], ",",colnames(wide_settle[c]))
    }
  }
}
```